\LARGE \bf Human-Robot Interaction: Tackling the AI Challenges

### 2014-05-02 paper.tex
Re-organization sect 2 <-> sect 3
- added a sub section on reasoning
- added a dedicated sub section on symbol grounding
- moved 'internal cognitive skills' to sect 3

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Implementing the Cognitive Skills}
\label{sec:impl}

Previous section provided an overview of our cognitive architecture for robots
involved in interactions with other social agents, as well as the knowledge
model we rely on.

We now introduce each of the components and give insights on their main
features.

\subsection{Internal Cognitive Processes}
\label{sect|intern}

\subsubsection{Symbolic Anchoring}

Regarding the anchoring question, this architecture is bidirectional. The
components we described provide a \textit{bottom-up} grounding process:
geometric reasoning and dialogue processing modules constantly build and push
new symbolic contents about the world to the knowledge base where it becomes
accessible to decisional layers. In parallel, the knowledge base relies on
reasoning in a \textit{top-down} way to produce new facts that may in return
trigger physical behaviours.

@@

Theory of Mind (originally defined in~\cite{Premack1978}) is the cognitive
ability that a subject possesses to represent the mental state of another
agent, possibly including knowledge that contradicts the subject's own model: for
example, a book can be at the same time \emph{visible} for myself, and \emph{not
visible} for you.

Children develop this skill, which is essential to understand others' perspectives during
interactions, around the age of three. It supposes the
ability to build, store and retrieve separate models of the knowledge of the
interactors.

Our knowledge base implements such a mechanism: when the robots infers a new
agent has been introduced in the knowledge base, it initializes a new,
independent, ontology for this agent. All the ontologies that are created share
the same common-sense knowledge, but rely on each agent's perspective for the
actual instantiation: the robot (geometrically) computes that the book is in
its own field of view, but not in the human one. The robot knowledge contains
the fact \stmt{book isVisible true} while the human model contains \stmt{book
isVisible false}.

@@

We rely however on this mechanism in certain cases: for instance, some modules
like the natural language processor use the {\tt short term} memory profile to
mark for a few seconds important concepts that are currently manipulated by the
robot as \emph{active concepts}. For example, if a human asks the robot: ``Give
me all red objects'', the human, the \concept{Give} action, and each red
objects that are found are successively marked as \emph{active concepts} by
inserting statements such as \stmt{human type ActiveConcept} in the short-term
memory (which can be considered, in this case, to be a working memory). We use
this feature to trace certain knowledge-related processes.

>>>>

\subsection{Symbol Grounding}

\emph{Grounding} (also called \emph{anchoring} when specifically referring to
the building of links between \emph{percepts} and \emph{physical
objects}~\cite{Coradeschi2003}) is the task consisting in building and
maintaining a bi-directional link between sub-symbolic representations (sensors
data, low-level actuation) and symbolic representations that can be manipulated
and reasoned about~\cite{Harnad1990}. This represents an important cognitive
skill, in particular in the human-robot interaction context: in that case, the
link that the robot has to established between its percepts and its symbol must
\emph{model} to a large extend the human representations in order to effectively
support communication.

\emph{Symbol grounding} relates the knowledge model to the perception and
actuation capabilities of the robot, and our architecture adopts here a
bidirectional approach.  The different components that we have described so far
provide a \textit{bottom-up} grounding process: geometric reasoning and dialogue
processing modules constantly build and push new symbolic contents about the
world to the knowledge base where it becomes accessible to decisional layers. In
parallel, the knowledge base relies on reasoning in a \textit{top-down} way to
produce new facts, \ie new beliefs that may lead to actual decision making and
physical actions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Cognitive Skills}
\label{sec:impl}

While the previous section provided an overview of our cognitive architecture,
as well as its associated knowledge model, this section discuss each of its main
``building blocks'' that we each consider to instantiate different cognitive
skills. Since each of these components have multiple connections to the others,
no obvious presentation order stands out, and the reader may have to rely on
Figure~\ref{fig|archi} as a guide.

We start by discussing the main \emph{internal} cognitive capabilities,
implemented in the knowledge base itself, and then discuss successively {\sc
Spark}, the situation assessment module, {\sc Dialogs}, the dialogue processor,
HATP, the symbolic task planner, and finally the main features of our execution
controllers.

\subsection{Internal Cognitive Skills}
\label{sect|intern}

We call \emph{internal} those cognitive capabilities that are tightly bound to
the knowledge model, and hence implemented directly inside {\sc Oro}, the knowledge base.
We present here three of them, \emph{reasoning}, \emph{theory of mind} modeling
and our (naive) approach to \emph{memory management}.

\subsubsection{Reasoning}

As mentioned in the previous section, we use the Pellet open-source reasoner to
reason on the knowledge base. This enables several standard inference services:
\emph{consistency checking}, \emph{concept satisfiability},
\emph{classification} and \emph{realisation} (computation of the most specific
classes that an individual belongs to). In case of logical inconsistency, the
reasoner can also provide \emph{explanations} that are important for debugging.

Besides those, {\sc Oro} server implements several algorithms to identify
similarities and differences between concepts (classes or
instances)~\cite{Ros2010b}: the \emph{Common Ancestors} algorithm, useful to
determine the most specific class(es) that include a given set of individuals;
the \emph{First Different Ancestors} algorithm that returns what can be
intuitively understood as the most generic types that \emph{differentiate} two
concepts; and \emph{clarification} and \emph{discrimination} algorithms that
play a key role in the process of \emph{interactive grounding} of the semantics
of concepts (we discuss this process at section~\ref{sect|com}). Clarification
and discrimination algorithms are based on what we call \emph{descriptors}, \ie
properties of individuals, either statically asserted in the common-sense
ontology, acquired by the robot through perception or proactive questioning of
the human partner, or derived from other reasoning algorithms like the
\emph{Common Ancestors} and \emph{Different Ancestors}. The
\emph{discrimination} algorithm consists then in looking for discriminants, \ie
descriptors that allow a maximum discrimination among a set of individuals.

@@

Theory of Mind (originally defined in~\cite{Premack1978}) is the cognitive
ability that allows a subject to represent the mental state of another
agent, possibly including knowledge that contradicts the subject's own model: for
example, a book can be at the same time \emph{visible} for myself, and \emph{not
visible} for you. Children develop this skill, which is essential to understand others' perspectives during
interactions, around the age of three.

From a robotics point of view, it supposes the ability to build, store and
retrieve separate models of the beliefs of the agents the robot interacts with.
Our knowledge base implements such a mechanism: when the robots recognises that
a new agent has been created in the knowledge base, it initializes a new,
independent, ontology for this agent. All the ontologies that are created share
the same common-sense knowledge, but rely on each agent's perspective for the
actual instantiation: the robot (geometrically) computes that the book is in its
own field of view, but not in the human one. The robot knowledge contains the
fact \stmt{book isVisible true} while the human model contains \stmt{book
isVisible false} (this computation, called \emph{perspective taking}, is
discussed in the next section).

@@

We rely however on this mechanism in certain cases: for instance, some modules
like the natural language processor use the {\tt short term} memory profile to
mark for a few seconds important concepts that are currently manipulated by the
robot as \emph{active concepts}: If a human asks the robot ``Give
me all red objects'', the human, the \concept{Give} action, and each red
objects that are found are successively marked as \emph{active concepts} by
inserting statements such as \stmt{human type ActiveConcept} in the short-term
memory (which can be considered, in this case, to be a working memory). We use
this feature to trace certain knowledge-related processes.
